---
title: "homework_9"
author: "Peter"
date: "`r Sys.Date()`"
output: html_document
---

HW # 9 — due before exam time Friday Apr 10 10 am 
Sec 8.9 Exercises #18, 19


### Exercise 18

a. \

```{r}

pedro_a <- 70
pedro_b <- 10
mia_a   <- 33.3
mia_b  <- 3.3

# Means
pedro_mean <- pedro_a / pedro_b
mia_mean   <- mia_a / mia_b

pedro_sd <- sqrt(pedro_a / pedro_b^2)
mia_sd   <- sqrt(mia_a / mia_b^2)

pedro_mean
pedro_sd
mia_mean
mia_sd
```

Mia expects more ER visits on average (10.09091 vs 7), but Pedro is more confident in his guess. His prior is much tighter (SD = 0.83666 vs 1.748671). \

b. \

```{r}
qgamma(c(0.05, 0.95), shape = 70, rate = 10)
qgamma(c(0.05, 0.95), shape = 33.3, rate = 3.3)
```

Pedro's 90% credible interval is (5.682967, 8.430648) and Mia's 90% credible interval is (7.396876, 13.128930). Mia's interval is wider than Pedro's, which makes sense because Mia's prior is less certain than Pedro's. \

c. \ 

Pedro would need to keep the ratio $\alpha/\beta = 7$ (same mean) but scale both parameters down to increase spread. For example, he could do $\alpha = 35, \beta = 5$: \

```{r}
qgamma(c(0.05, 0.95), shape = 35, rate = 5)
```

This keeps the mean of $7$, but his 90% CI is now (5.173928, 9.053123), which reflects his uncertainty. \

d. \

d. Mia can keep $\alpha = 33.3$ fixed and only adjust $\beta$ to shift the mean up by 3. New $\beta = 33.3/(10.09091 + 3) = 2.54375$: \

```{r}
beta <- 33.3/(10.09091 + 3)
qgamma(c(0.05, 0.95), shape = 33.3, rate = beta)
```


This keeps the same degree of confidence, but her 90% CI is now (9.595948, 17.032126), which is wider and shifted up from her original (7.396876, 13.128930). \



### Exercise 19

$$\lambda \mid \text{data} \sim \text{Gamma}(\alpha + \sum{y}, \beta + n)$$

a. \

```{r}
y <- c(8, 6, 6, 9, 8, 9, 7)
alpha_post <- 70 + sum(y)
beta_post  <- 10 + length(y)
alpha_post/beta_post

qgamma(c(0.025, 0.975), shape = alpha_post, rate = beta_post)
```

The posterior is $\text{Gamma}(123, 17)$ with mean $7.235294$. The $95\%$ credible interval for $\lambda$ is $(6.013237, 8.568716)$. So, we believe there is a 95% chance the true average number of ER visits per hour is between 6.013237 and 8.568716. \

b. \

```{r}
pgamma(6, shape = alpha_post, rate = beta_post)
```

The posterior probability that $\lambda \leq 6$ is only $0.02368644$. This means there is less than a $2.4\%$ chance that the true average number of ER visits is at most 6, given the data. The administrator's claim is very unlikely to be true. \


c. \

```{r}
set.seed(6549534)
S <- 100000

week_tot <- replicate(S, {
  lambda_sim <- rgamma(1, shape = alpha_post, rate = beta_post)
  sum(rpois(7, lambda_sim))
})

mean(week_tot)
quantile(week_tot, c(0.025, 0.975))
```

Using $100000$ simulations, the predicted total number of ER visits for the next week has a mean of $50.61202$, with a $95\%$ posterior predictive interval of $(35, 68)$. \










