---
title: "homework_4"
author: "Peter"
date: "`r Sys.Date()`"
output: html_document
---


Sec 4.6 Exercises # 3, 5, 6, 9, 12, 15, 18, 23, 29

### Problem 3

a. 

$$
P(X=x) = \newline

X=1: \frac{2}{32} \newline
X=2: \frac{8}{32} \newline
X=3: \frac{12}{32} \newline
X=4: \frac{8}{32} \newline
X=5: \frac{2}{32} \newline
$$

b.

```{r}
x <- 1:5
px <- c(2/32, 8/32, 12/32, 8/32, 2/32)
barplot(px, names.arg = x, xlab = "X", ylab = "P(X = x)")
```

c.
The most likely number of runs is the mode, which is $X=3$. \

d.

$P(X \leq 2) = P(X=1) + P(X=2) = \frac{10}{32}$.



### Problem 5

a.

$$
P(X=x) = \newline

X=1: \frac{1}{6} \newline
X=2: \frac{1}{6} \newline
X=3: \frac{1}{6} \newline
X=4: \frac{1}{4} \newline
X=5: \frac{1}{4} \newline
$$

b.

$P(X \geq 2) = 1 - P(X=1) = 1 - \frac{1}{6}=\frac{5}{6}$

c.

$\mu = \sum x \cdot P(X=x) = \frac{13}{4} = 3.25$

$E[X^2] = \frac{14}{6} + \frac{41}{4} = \frac{151}{12}$

$\sigma^2 = \frac{151}{12} - 3.25^2 = 2.020833$

$\sigma = 1.42156$


### Problem 6

a.

```{r}
x <- 4:24
px <- c(.001,.003,.008,.016,.027,.044,.062,.080,.097,.108,.113,.108,.096,.080,.062,.043,.027,.015,.008,.003,.001)

y <- seq(4, 24, by = 2)
py <- c(.028,.056,.083,.111,.139,.167,.139,.111,.083,.056,.028)

mu_x <- sum(x * px)
sd_x <- sqrt(sum(x^2 * px) - mu_x^2)

mu_y <- sum(y * py)
sd_y <- sqrt(sum(y^2 * py) - mu_y^2)

cat("X: mean =", mu_x, ", sd =", sd_x, "\n")
cat("Y: mean =", mu_y, ", sd =", sd_y, "\n")
```

$\mu_X = 14.014$
$\sigma_X = 3.371617$

$\mu_Y = 14.014$
$\sigma_Y = 4.817863$


b.

```{r}
plot(x, px, type = "h", lwd = 2, col = "blue", xlab = "Value", ylab = "Probability")
points(y + 0.2, py, type = "h", lwd = 2, col = "red")
legend("topright", legend = c("X", "Y"), col = c("blue", "red"), lwd = 2)
```


c.

They are symmetric, having the same mean, though Y has a larger standard deviation since it only takes even values. \
The means match, but doubling increases variability too much and makes odd totals impossible. Four independent dice cancel out extremes better, giving a tighter distribution.


### Problem 9

a.

$$
P(Y=y) = \newline

Y=-4: \frac{1}{16} \newline
Y=-2: \frac{4}{16} \newline
Y=0: \frac{6}{16} \newline
Y=2: \frac{4}{16} \newline
Y=4: \frac{1}{16} \newline

$$

b. 

$P(|Y| \geq 2) = P(Y = -4) + P(Y = -2) + P(Y = 2) + P(Y = 4) = \frac{10}{16}$

c.

$$
P(Y=y) = \newline

Y=-4: (0.5)(0.3)(0.3)(0.3) = 0.0135 \newline
Y=-2: (0.0315)+(0.0315)+(0.0525)+(0.0525) = 0.168 \newline 
Y=0: (0.0735)+(0.1225)+(0.1225)+(0.1225)+(0.1225)+(0.0735)= 0.637 \newline
Y=2: (0.0315)+(0.0315)+(0.0525)+(0.0525) = 0.168 \newline
Y=4: (0.5)(0.3)(0.3)(0.3) = 0.0135 \newline

$$


d.


$\mu_a = 0$
$\sigma_a = \sqrt{16(\frac{1}{16}) + 4(\frac{4}{16}) + 4(\frac{4}{16}) + 16(\frac{1}{16})} = 2$

$\mu_c = 0$
$\sigma_c = \sqrt{16(0.0135) + 4(0.168) + 4(0.168) + 16(0.0135)} = 1.333$

Both distributions have the same mean of 0, but the gravitational pull reduces the standard deviation from 2 to 1.33, so the walker stays much closer to home on average.



### Problem 12


### Problem 15


### Problem 18


### Problem 23


### Problem 29





